---
title: "GroupProject"
author: "Jun Hernandez, Victoria Rios, "
date: "May 6, 2020"
output: pdf_document
---

```{r}
#install.packages("read_xl")
#install.packages("tidyverse") # uncomment to install tidyverse and make sure you are on up to date version
#install.packages("rcompanion")
#install.packages("DescTools")
```

```{r}
library(tidyverse)
library(readxl)
library(rcompanion)
```

```{r}
# read in xlsx data set
df <- read_excel("/Users/junhernandez/Documents/A\ -\ Current\ Quarter/Stat\ 183/Group\ Project/all\ responses\ data\ team\ 1.xlsx") # change for your computer

#View(df)

head(df)
# test
```

**Q1**

**Q2**

```{r q2}
df_fixed <- df[-c(1), ]
glimpse(df_fixed)

df_fixed %>%
  select(
    info = `There are many ways to get health information on campus. Which way would you most prefer?`,
    other = ...36,
    gender = 'What is your gender?') %>%
  filter(!(is.na(gender))) %>% #removing responses without gender
  filter(gender == 'Male'| gender == 'Female') 
```


```{r}
# Remove first row
df_fixed <- df[-c(1), ]
# subset dataset for #4
q2_df <- df_fixed %>%
  select(	
   info = `There are many ways to get health information on campus. Which way would you most prefer?`,
   gender =`What is your gender?`) %>%
  filter(!(is.na(gender))) %>%
  filter(gender == 'Male'| gender == 'Female') %>%   # only focused on male and female data
  filter(!(is.na(info)))   # 16 NAs filtered in culture

# set vars as factor
q2_df$info <- as.factor(q2_df$info)
q2_df$gender <- as.factor(q2_df$gender)
  
q2_df


```

```{r}
# based on count
q2_df %>% ggplot(mapping = aes(x = info)) +
  geom_bar(aes(fill = info)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "There are many ways to get health information on campus. Which way would you most prefer?") +
  theme(legend.position="none")
# add sub title for No. of participants

# based on count
q2_df %>% ggplot(mapping = aes(x = info)) +
  geom_bar(aes(fill = gender)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "There are many ways to get health information on campus. Which way would you most prefer?") 
```

```{r}
# based on proportion
q2_df %>% filter(gender == 'Female') %>%
   ggplot(mapping = aes(x = info, 
                        y = prop.table(stat(count)),
                        fill = info,
                        label = scales::percent(prop.table(stat(count))))) + theme(axis.text.x=element_text(angle=90,hjust=1)) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Female") +
   theme(legend.position="none")
```

```{r}
# based on proportion
q2_df %>% filter(gender == 'Male') %>%
   ggplot(mapping = aes(x = info, 
                        y = prop.table(stat(count)),
                        fill = info,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Male") +
   theme(legend.position="none")

```
**Chi-squared test**

```{r}
#str(q2_df)
res <- chisq.test(q2_df$gender, q2_df$info)
# Access to the GoF values
res$expected
res$observed
res$stdres
# output
res
```

```{r}
tbl <- table(q2_df$info, q2_df$gender)
chisq.test(tbl)

pairwiseNominalIndependence(tbl,
                            fisher = FALSE,
                            gtest  = FALSE,
                            chisq  = TRUE ,
                            method = "bonferroni")
```




#5 using data from column 22  
In which of the following categories do you place yourself?  

```{r q5 data subset}
# get rid of first row
# commenting this out for now
df_fixed <- df[-c(1), ]

# subset dataset for #5
victoria_df <- df_fixed %>%
  select(	
   lifestyle = `In which of the following categories do you place yourself?`,
   gender =`What is your gender?`) %>%
  filter(!(is.na(gender))) %>% #removing responses without gender
  filter(gender == 'Male'| gender == 'Female') %>%   # removing 29 'other'
  filter(!(is.na(lifestyle)))   # 14 NAs filtered in lifestyle (people who didn't respond)

# set vars as factor
victoria_df$lifestyle <- as.factor(victoria_df$lifestyle)
victoria_df$gender <- as.factor(victoria_df$gender)
  
glimpse(victoria_df)
```

```{r q5 overview}
# based on count
victoria_df %>%
ggplot(mapping = aes(x = lifestyle,
label = stat(count))) +
geom_bar(aes(fill = lifestyle)) +
labs(x = element_blank(),
  y = "Number of responses",
     title = "Overall Summary",
subtitle = "No. of Participants: 1,311") +
theme(legend.position="none",
axis.text.x = element_text( hjust=0.5)) + # center align the axis labels
geom_text(stat = 'count',
position = position_dodge(.9),
vjust = -0.5,
size = 3) +
# my variable names are really long
scale_x_discrete(
labels = c("I am not interested \nin pursuing a \nhealthier lifestyle",
"I am thinking about \nchanging some of my \nhealth behaviors",
"I am planning on \nmaking a health \nbehavior change within \nthe next 30 days",
"I have made some \nhealth behavior \nchanges within \nthe last 6 months",
"I have been living \na healthy lifestyle \nfor at least 6 months")) 
  ggsave("q5overallcount.png", width = 6, height = 4)
```

```{r option1: facet wrap}
# based on proportion
victoria_df %>%
#  filter(gender == 'Female') %>%
ggplot(mapping = aes(x = lifestyle,
y = prop.table(stat(count)),
fill = lifestyle,
label = scales::percent(prop.table(stat(count))))) +
geom_bar(position = "dodge") +
facet_wrap(~gender, nrow = 1) +
geom_text(stat = 'count',
position = position_dodge(.9),
vjust = -0.5,
size = 3) +
scale_y_continuous(labels = scales::percent) +
labs(x = "Survey Response",
y = "Proportion of responses",
title = "Response by Gender") +
theme(legend.position="right",
legend.title = element_blank(),
legend.text = element_text(lineheight = .8),
axis.text.x = element_blank(),
axis.ticks.x = element_blank()) +
scale_fill_discrete( 
labels = c("I am not interested \nin pursuing a \nhealthier lifestyle",
"I am thinking about \nchanging some of my \nhealth behaviors",
"I am planning on \nmaking a health \nbehavior change within \nthe next 30 days",
"I have made some \nhealth behavior \nchanges within \nthe last 6 months",
"I have been living \na healthy lifestyle \nfor at least 6 months"))
```

```{r option2}
# based on proportion
victoria_df %>%
#  filter(gender == 'Male') %>%
ggplot(mapping = aes(x = lifestyle,
                     y = prop.table(stat(count)),
                     fill = gender,
                     label = scales::percent(prop.table(stat(count))))) +
  geom_bar(position = "dodge") +
  geom_text(stat = 'count',
            position = position_dodge(.9),
            vjust = -0.5,
            size = 3) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = element_blank(),
       y = "Proportion of responses",
       title = "Stratified Analysis by Gender",
       subtitle = "Q5: 902 females & 409 males") +
  theme(legend.position="right",
        axis.text.x = element_text(hjust=0.5)) + # center align the axis text)
  # my variable names are really long
  scale_x_discrete(
labels = c("I am not interested \nin pursuing a \nhealthier lifestyle",
"I am thinking about \nchanging some of my \nhealth behaviors",
"I am planning on \nmaking a health \nbehavior change within \nthe next 30 days",
"I have made some \nhealth behavior \nchanges within \nthe last 6 months",
"I have been living \na healthy lifestyle \nfor at least 6 months"))

ggsave("q5gender.png", width = 6, height = 4)
```

```{r chisquared test}
victoria_res <- chisq.test(victoria_df$gender, victoria_df$lifestyle)
# Access to the GoF values
victoria_res$expected
victoria_res$observed
victoria_res$stdres
# output
victoria_res

victoria_tbl <- table(victoria_df$gender, victoria_df$lifestyle)
victoria_tbl

chisq.test(victoria_tbl)

pairwiseNominalIndependence(victoria_tbl,
                            fisher = FALSE,
                            gtest  = FALSE,
                            chisq  = TRUE,
                            method = "bonferroni")

```

Frequency distributions are significantly different between males and females (chi-square test p-value = .0013). More women (24%) than men(9%) state that they have been living a healthy lifestyle for at least 6 months.  





#7 using # 23 data(gender - m/f)

```{r q7}
# get rid of first row
df_fixed <- df[-c(1), ]
# subset dataset for #7
jun_df <- df_fixed %>%
  select(	
   culture = `UCR has a culture and environment that promotes and supports health and well-being for its faculty, staff and students.`,
   gender =`What is your gender?`) %>%
  filter(!(is.na(gender))) %>%
  filter(gender == 'Male'| gender == 'Female') %>%   # only focused on male and female data
  filter(!(is.na(culture)))   # 16 NAs filtered in culture

# set vars as factor
jun_df$culture <- as.factor(jun_df$culture)
jun_df$gender <- as.factor(jun_df$gender)
  
jun_df

#unique(jun_df$gender)
#unique(jun_df$culture)
```

```{r q7_counts}
# Plot the data for #7 
# based on count
jun_df %>% ggplot(mapping = aes(x = culture,
                  label = stat(count))) +
  geom_bar(aes(fill = culture)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "UCR has a culture and environment that promotes and supports health 
       and well-being for its faculty, staff and students",
       subtitle = "(N = 1309)") +
  theme(legend.position="none",
        axis.text.x = element_text( hjust=0.5)) + # center align the axis labels
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) 

# Plot the data for #7 with gender
# based on count
jun_df %>% ggplot(mapping = aes(x = culture,
                  label = stat(count))) +
  geom_bar(aes(fill = gender)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "UCR has a culture and environment that promotes and supports health 
       and well-being for its faculty, staff and students",
       subtitle = "(N = 1309)") +
  theme(axis.text.x = element_text( hjust=0.5)) + # center align the axis labels
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) 
```

```{r q7_female}
# based on proportion for female only responses
jun_df %>% filter(gender == 'Female') %>%
   ggplot(mapping = aes(x = culture, 
                        y = prop.table(stat(count)),
                        fill = culture,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Female",
       subtitle = "(N = 904)") +
   theme(legend.position="none")
```

```{r q7_male}
# based on proportion, Male responses only
jun_df %>% filter(gender == 'Male') %>%
   ggplot(mapping = aes(x = culture, 
                        y = prop.table(stat(count)),
                        fill = culture,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Male",
       subtitle = "(N = 405)") +
   theme(legend.position="none")

```

```{r q7_gender}
# based on proportion, stratified by gender
jun_df %>%
   ggplot(mapping = aes(x = culture, 
                        y = prop.table(stat(count)),
                        fill = gender,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Stratified analysis by Gender",
       subtitle = "(N = 1309)") 
```

**Chi-squared test**

```{r}
#str(jun_df)
res <- chisq.test(jun_df$gender, jun_df$culture)
# Access to the GoF values
res$expected
res$observed
res$stdres
# output
res
```
The frequency distributions are significantly different between females and males (chi-squared test p-value = 0.0001). More women agree and strongly agree that UCR has a culture and environment that promotes and supports health and well-being for its faculty, staff and students, while more men disagree, strongly disagree, and are neutral to the statement. 


**Question 11**

```{r}
# Remove first row
df_fixed <- df[-c(1), ]
# subset dataset for #11
q11_df <- df_fixed %>%
  select(	
   diet = `A healthful well-balanced diet consists of a daily intake from all food groups with an emphasis on fresh produce, lean protein, whole grains and heart-healthy unsaturated oils; keeping sugar, sodium and saturated fat to a minimum.Based on the aforementioned parameters, how would you rate your own dietary intake?`,
   gender =`What is your gender?`) %>%
  filter(!(is.na(gender))) %>%
  filter(gender == 'Male'| gender == 'Female') %>%   # only focused on male and female data
  filter(!(is.na(diet)))   # 16 NAs filtered in culture

# set vars as factor
q11_df$diet <- as.factor(q11_df$diet)
q11_df$gender <- as.factor(q11_df$gender)
  
q11_df

```

```{r}
# based on count
q11_df %>% ggplot(mapping = aes(x = diet)) +
  geom_bar(aes(fill = diet)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "How would you rate your own dietary intake?") +
  theme(legend.position="none")
# add sub title for No. of participants

# based on count
q11_df %>% ggplot(mapping = aes(x = diet)) +
  geom_bar(aes(fill = gender)) +
  labs(x = "Survey Response", 
       y = "Number of responses", 
       title = "How would you rate your own dietary intake?")
```

```{r}
# based on proportion
q11_df %>% filter(gender == 'Female') %>%
   ggplot(mapping = aes(x = diet, 
                        y = prop.table(stat(count)),
                        fill = diet,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Female") +
   theme(legend.position="none")
```

```{r}
# based on proportion
q11_df %>% filter(gender == 'Male') %>%
   ggplot(mapping = aes(x = diet, 
                        y = prop.table(stat(count)),
                        fill = diet,
                        label = scales::percent(prop.table(stat(count))))) +
   geom_bar(position = "dodge") + 
   geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 3) + 
   scale_y_continuous(labels = scales::percent) +
   labs(x = "Survey Response", 
       y = "Proportion of responses", 
       title = "Male") +
   theme(legend.position="none")

```
**Chi-squared test**

```{r}
#str(q4_df)
res <- chisq.test(q11_df$gender, q11_df$diet)
# Access to the GoF values
res$expected
res$observed
res$stdres
# output
res
```

```{r}
tbl <- table(q11_df$diet, q11_df$gender)
chisq.test(tbl)

pairwiseNominalIndependence(tbl,
                            fisher = FALSE,
                            gtest  = FALSE,
                            chisq  = TRUE ,
                            method = "bonferroni")
```



